Environmental warming increases the importance of high-turnover energy channels in stream food webs

### 2022-12-01 doc/ms/Junker_temp-energy-flux_ms.Rmd
updated manuscript version 2

---
title: "Environmental warming increases the absolute and relative importance of high-turnover energy channels in stream food webs"
author:
  - James R. Junker:
      email: jrjunker@mtu.edu
      institute: [msu, glrc]
      correspondence: true
  - name: Wyatt F. Cross
    institute: msu
  - name: James M. Hood
    institute: osu
  - name: Jonathan P. Benstead
    institute: bama
  - name: Alex D. Huryn
    institute: bama
  - name: Daniel Nelson
    institute: usu
  - name: Jón S. Ólafsson
    institute: veidi
  - name: Gísli M. Gíslason
    institute: reyk
institute:
  - msu: Department of Ecology, Montana State University, Bozeman, MT 59717, USA
  - glrc: Current `address:` Great Lakes Research Center, Michigan Technological University, Houghton, MI 49931, USA
  - osu: Department of Evolution, Ecology, and Organismal Biology, The Ohio State University, Translational Data Analytics Institute, The Aquatic Ecology Laboratory, Columbus, OH 43212, USA
  - bama: Department of Biological Sciences, University of Alabama, Tuscaloosa, AL  35487, USA
  - usu: National Aquatic Monitoring Center, Department of Watershed Sciences, Utah State University, Logan, UT 84322, USA
  - veidi: Institute of Marine and Freshwater Fisheries, Reykjavik, Iceland
  - reyk: Univeristy of Iceland, Institute of Life and Environmental Sciences, Reykjavik, Iceland
output:
   word_document:
      reference_docx: working_docx_template.docx
      toc: no
      pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
      - '--filter=pandoc-crossref.exe'
   pdf_document:
     keep_tex: true
   html_document: default
   pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
      - '--filter=pandoc-crossref.exe'
header-includes:
- \usepackage{lineno}
- \usepackage{amsmath}
- \numberwithin{equation}
- \usepackage{indentfirst}
- \linenumbers
indent: true
linestretch: 1
bibliography: refs.bib
link-citations: no
link-color: grey
csl: ecology.csl
---

@@

Warming temperatures are altering communities and trophic networks globally. While the influence of warming on food webs is often context-dependent, increasing temperatures are predicted to influence communities in two fundamental ways: 1) by reducing average body size and 2) by increasing individual metabolic rates. These warming-induced changes have the potential to influence the distribution of food web fluxes, food web stability, and the importance of environmental and biological factors in community assembly. Here, we quantified patterns and the relative distribution of organic matter fluxes through stream food webs spanning a natural ~25$^\circ$C temperature gradient. We then related these patterns to species- and community-level trait distributions (body size, *M* and biomass turnover, *P:B*) within and across ecosystems. We predicted that 1) communities in warmer streams would skew towards smaller *M* and higher *P:B* and 2) organic matter fluxes within warmer communities would increasingly skew towards smaller, higher *P:B* populations. Across the temperature gradient, communities in warmer streams were indeed composed of smaller-bodied and higher biomass-turnover populations on average. Additionally, material fluxes within warmer streams were increasingly skewed towards high-turnover populations, demonstrating that warming can restructure organic matter fluxes in both an absolute and relative sense. With warming, the relative distribution of organic matter fluxes also appeared to be increasingly 'non-random', suggesting the potential for stronger selection for these traits with increasing temperature. Our study suggests that a warming world will be 'smaller and faster', and that warming has the potential to shape pathways of energy and material flows in food webs in ways that may be predictable.

@@

<!--- warming and effects in food webs --->
<!--- Speeding up of food web: Species levels--->
Increasing global temperatures can influence the provision and maintenance of ecosystem services by modifying the network of species interactions that underpin ecosystem functions [@brose2012; @woodward2010; @deruiter1995; @thompson2012]. The effects of warming permeate across levels of biological organization from its control on individual metabolic rates [@gillooly2001; @brown2004] and biological activity (e.g., attack rate, handling time, growth rates, etc.; @dell2014, to broad-scale shifts in community assembly and structure [@saito2021; @gibert2019a; @nelson2017a]. Across global climate gradients, these temperature-induced changes have the potential to mediate food web stability [@baiser2019] by altering the acquisition and allocation of resources [@zhang2017] and the magnitudes and relative distribution of energy fluxes in food webs [@may1972; @mccann1998; @barnes2018].

Understanding how warming alters energy fluxes in food webs requires information about how temperature modifies connections between ecosystem structure (i.e. biodiversity) and function. Such connections hinge upon how warming influences functional trait distributions and how these functional traits translate to energy demands [@loreau2001; @norberg2001; @norberg2004]. Additionally, warming may modify both deterministic [e.g., environmental/niche filtering, @whittaker1962] and stochastic [e.g., neutral theory, @hubbell2001] community assembly processes [@saito2021], leading to many possible relationships between warming and species' abundance and trait distributions. For example, although temperature appears to have minimal control on species richness across taxonomic groups [e.g., @bastazini2021], warming may have profound effects on the dominance structure (i.e. evenness) of ecological communities, often by reducing community evenness and favoring a few well-adapted species [@hillebrand2008]. This strong environmental filtering is likely to skew trait distributions, to some extent, in natural communities [@therriault1999]. However, the relative distribution of traits can be modified by additional processes beyond environmental filtering that alter species relative abundance distributions [e.g., species interactions, @therriault1999; demographic stochasticity, @hubbell2001]. These additional processes can exaggerate or counter any skew in trait distribution imparted through environmental filtering and may modify the relationship between species' traits and the absolute and relative energy demands in food webs.

<!--- Connect traits to the 'speed' of ecosystem energy fluxes--->
Body size is a fundamental trait that is strongly influenced by temperature and has great potential to influence energy flux [@atkinson1994; @gardner2011; @daufresne2009]. Body size reductions arise from multiple mechanisms acting at different levels of organization, including the increased relative abundance of smaller species within warmer communities [@bergmann1848] or smaller individuals in warmer populations [@james1970], or reduced absolute body size of warmer individuals [@atkinson1994]. These temperature-size relationships are likely to influence the distribution of body sizes both across and within ecosystems, such that warmer communities should contain smaller species and populations, on average, and smaller organisms should be more dominant within a community. These changes can have important implications for ecosystems pattern and processes that control energy and material fluxes in ecosystems because body size is a strong determinant of species life history patterns [@peters1983; @zeuss2017; @altermatt2010; @nelson2020b] and, developmental [@angilletta2004] or metabolic rates [@brown2004; @gillooly2001].

<!--- Write the main goals of the paragraph--->
Metabolic rate, in addition to being controlled by body size, is another key trait that is influence by temperature and tied to variation in energy flux through food webs [@gillooly2001]. Warming can influence metabolic rates indirectly through reductions in body size, as well as directly through its effects on subcellular kinetics [@bideault2019; @osmond2017]. These processes have been shown to modify ecosystem patterns through changes in population abundance [@bernhardt2018], consumer-resource interactions [@bideault2019], and food web structure [@gibert2019a]. Importantly, the kinetic effects of temperature are tied to many biological processes [@dell2014], among them growth rate [@gillooly2001], developmental rate [@zuo2012; @nelson2020b], voltinism [@zeuss2017], and population turnover rate [@huryn2007; @brown2004]. Taken together, these changes suggest that warming should lead to a smaller, faster world in which ecosystem processes 'speed up' through the effects of smaller body size and higher turnover rates, with a potentially strong imprint on food web structure and energy fluxes [@gibert2019a].

We quantified patterns of organic matter flux in stream food webs across a natural temperature gradient (\~5--28$^\circ$C) in southwestern Iceland. Previous research in these streams has shown a strong positive effect of warming on primary production both among streams [@demars2011; @padfield2017] and within streams seasonally [@ogorman2012; @hood2018]. Invertebrates in these streams rely on autochthonous resources [@ogorman2012; @nelson2020], thus the dynamics of primary production have a strong control on consumer energy demand [@junker2020]. We, therefore, predicted that annual organic matter fluxes to consumers would increase with stream temperature mirroring patterns in resource availability and consumer energy demand. We also hypothesized that temperature would act as a principle environmental filter on community assembly and organic matter fluxes by favoring 'fast' life-history traits. Specifically, we predicted that warming temperatures among streams would lead to a decrease in average body size and an increase in average biomass turnover (i.e., *P:B* ratio) of populations. We also predicted that *within* communities, organic matter fluxes would be dominated by small-bodied and high *P:B* taxa at higher temperatures and that these patterns would not arise by random change [i.e., a 'selection effect' *sensu* @huston1997]. Our results should help refine general predictions about how warming, and its influence on key traits, is likely to shape energy flux through animal food webs.

@@

We studied six streams within the Hengill geothermal field of southwestern Iceland (64$^\circ$ 03'N 021$^\circ$ 18'W) that varied in mean annual temperature. Hengill is
characterized by indirect geothermal heating of groundwater [@arnason1969], leading to natural variability in stream temperatures (4.5--54.0 $^\circ$C), but similar solute
chemistries [@friberg2009]. These conditions create a "natural laboratory" for isolating the effects of temperature on ecosystem processes [@ogorman2014; @nelson2017a]. We selected streams to maximize the temperature range, while minimizing differences in the structural aspects of the primary producer community [@junker2020a]. In each stream, we measured temperature and water depth every 15 min from July 2010 through August 2012 (U20-001-01 water-level logger, Onset Computer Corp. Pocasset, MA, USA). Light availability in the watershed was measured every 15 min from atmospheric stations (HOBO pendant temperature/light UA-002-64, Onset Computer Corp. Pocasset, MA, USA).

@@

We sampled macroinvertebrate communities approximately monthly in six streams, four from July 2011 to August 2012 and two from October 2010 to October 2011 (*n* = 6 streams). The two streams sampled from 2010--2011 were part of a separate warming manipulation during the reference (i.e., non-manipulated) year [@nelson2017; @nelson2017a]. Inter-annual comparisons of primary and secondary production in previous studies showed minimal differences among years in unmanipulated streams, suggesting that combining data from different years would not significantly bias our results [@nelson2017; @hood2018]. We collected five Surber samples (0.023 m^2^, 250 $\mu$m mesh) from randomly selected locations within each stream. Within the sampler, inorganic substrates were disturbed to \~10 cm depth and invertebrates and
organic matter were removed from stones with a brush. Samples were then preserved with 5% formaldehyde until laboratory analysis. In the laboratory, we split samples into coarse
(\>1 mm) and fine (\>250 $\mu$m -- \<1 mm) fractions using nested sieves and then removed invertebrates from each fraction under a dissecting microscope (10--15 x magnification).
For particularly large samples, fine fractions were sub-sampled (1/2--1/16th) using a modified Folsom plankton splitter prior to removal of invertebrates. Subsamples were scaled to the rest of the sample assuming similar abundance and body size distributions. Macroinvertebrates were identified to the lowest practical taxonomic level (usually genus) with taxonomic keys [@peterson1977; @merritt2008; @andersen2013]. Taxon-specific abundance and biomass were scaled to a per meter basis by dividing by the Surber sampler area.

@@

Daily secondary production of invertebrate taxa was calculated using the instantaneous growth rate method [IGR, @benke2017]. Growth rates were determined using taxon-appropriate
approaches described in Junker and others [-@junker2020]. Briefly, growth rates of common taxa (e.g., Chironomidae spp., *Radix balthica*, etc.) were determined using *in situ* chambers [\ *sensu* @huryn1986].
Multiple individuals (*n* = 5--15) within small size categories (\~1 mm length range) were photographed next to a field micrometer, placed in the stream within pre-conditioned
chambers for 7--15 days, and removed and photographed. Individual lengths were measured from field pictures using image analysis software [@schindelin2012], and body
lengths were converted to mass (mg ash-free dry mass [AFDM]) using published length-mass regressions [@benke1999; @ogorman2012; @hannesdottir2013]. Growth rates (*g*, d^-1^) were
calculated by the changes in mean body size (*M*) over a given time interval (*t*) with the following equation:

@@

Variability in growth rates was estimated by bootstrapping through repeated resampling of individual lengths with replacement (*n* = 1000). For taxa that exhibit synchronous growth and development (e.g., Simuliidae spp., some Chironomidae spp., etc.), we examined temporal changes in length-frequency distributions and calculated growth rates and uncertainty using a bootstrap technique similar to that described in Benke and Huryn [-@benke2017]. Individual lengths were converted to mass (mg AFDM) using published length-mass regression cited above, and size-frequency histograms were visually inspected for directional changes in body size through time. For each date, size-frequency distributions were resampled with replacement and growth rates estimated from equation 1. We prevented the calculation of negative growth rates by requiring $M_{t + \Delta t}$ \> $M_t$. If this condition was not met after 1e^4^ resamplings, a minimum growth rate of 0.001 was used. To estimate growth rates of taxa for which growth could not be estimated empirically, we developed stream-specific growth rate models by constructing multivariate linear regressions of empirical growth rates against body size and temperature. To estimate uncertainty in production of each taxon, we used a bootstrapping technique that resampled measured growth rates, in addition to abundance and size distributions from individual samples. For each iteration, size-specific growth rates were multiplied by mean
interval biomass for each size class and the number of days between sample dates to estimate size class-specific production. For each interval, size classes were summed for each taxon to calculate total population-level production. Intervals were summed to estimate annual secondary production (g AFDM m^-2^ y^-1^).

@@

Organic matter fluxes through the community were calculated using the trophic basis of production method [TBP, @benke1980]. Taxon-specific secondary production estimates were combined with diet proportions (see Supplemental Materials), diet-specific assimilation efficiencies, *AE~i~*, and net production efficiencies, *NPE* [e.g., @mccullough1975], to estimate consumption of organic matter. For each food
category, *i*, diet proportions were multiplied by the gross growth efficiency ($GGE_{i} = AE_{i} * NPE$) to calculate the relative production attributable to each food
category. The relative production from each food type was then multiplied by the interval-level production and finally divided by $GGE_{i}$ to estimate consumption of organic matter from each food category by each taxon [@benke1980]. Consumption was calculated for each taxon across sampling intervals (typically \~1 month). Total interval consumption was calculated by summing across all taxa, while annual consumption was calculated by summing across all taxa and intervals. Variability in consumption estimates was estimated through a Monte Carlo approach, wherein bootstrapped vectors of secondary production for each taxon (see *Secondary production* methods above) were resampled and consumption estimated with the TBP method using modeled diet proportions (see *Diet analysis* in Supplemental Materials), diet-specific assimilation efficiencies, and net production efficiency. Variability in *AE~i~* was incorporated by resampling values from beta distributions fit to median and 2.5% and 97.5% percentiles for each diet item: diatoms = 0.30 (95% percentile interval (PI): 0.24-0.36), filamentous and green algae = 0.30 (95% PI: 0.24-0.36), cyanobacteria = 0.10 (95% PI: 0.08-0.12), amorphous detritus = 0.10 (95% PI: 0.08-0.12), vascular and non-vascular plants (bryophytes) = 0.1 (95% PI: 0.08-0.12), and animal material = 0.7 (95% PI: 0.56-0.84)[@welch1968; @cross2007; @cross2011; @benke1980; @benke1997]. Variability in *NPE* was incorporated by resampling values from an assumed beta distribution with median *NPE* = 0.45 (95% PI = 0.4-0.5). Beta distributions were fit using the 'get.beta.par()' function within the *rriskDistributions* package [@belgorodski2017].

@@

### Evenness Among taxa

To visualize and quantify the evenness of OM fluxes among taxa within each stream, we constructed Lorenz curves [@lorenz1905] on rank-ordered OM fluxes, such that in a community with $S$ species and the relative OM flux of species *i*, $p_i$, is ordered $p_1 \leq p_2 \leq ... p_S$. The Lorenz curve shows how a value, in this case OM flux, accumulates with increasing cumulative proportion of taxa. In a community with perfectly equal distribution of OM flux among taxa, the Lorenz curve is simply a straight diagonal line. Deviation from perfect equality was calculated as the Gini coefficient [@gini1921], normalized for differences in $S$ among streams, $G^*$ [@solomon1975; @chao2019]:

@@

where $G^*$ represents an index of relative evenness of OM fluxes bounded between zero and one; a value of one represents a community with equal proportion of total community OM flux for all species ($1/S$), and a value of zero represents a community in which the total OM flux is attributed to a single taxa.

@@

We predicted that warming would favor taxa with smaller body size and higher population turnover and therefore OM fluxes would be skewed towards small body size (*M*) and higher *P:B* across and within communities. To assess the potential effects of environmental filtering across communities, we used bootstrapped linear regressions to quantify the relationship between mean annual temperature ($^\circ$C) and mean population body size ($\overline{M}$) and biomass turnover ($\overline{P:B}$) of each stream community. Here, `r temperature_stats[['n_boot']]` values of population *M* or *P:B* were resampled with replacement from population secondary production vectors (see *Secondary Production* above) for each taxa within each stream. The mean of all populations within a stream was calculated and a linear model was fit between *log~e~*-transformed $\overline{M}$ or $\overline{P:B}$ and mean annual temperature. Response variables were transformed to meet the assumption of normally distributed residual variation.

To quantify how the distribution of organic matter fluxes were modified *within* a community and whether this modification was related to temperature, we assessed the extent to which the relative OM fluxes among taxa were skewed towards populations with lower or higher relative *M* or *P:B*. To do this, we ordered taxa based on within-stream rankings of annual population traits (i.e., *M*, *P:B*) and then calculated a measure of skewness, $Sk_{flux}$, based on quartiles of the distribution of OM fluxes in relation to taxa traits as:

@@

where, $f_{Qx}$, is the cumulative flux at some quantile, $Qx$, of the community trait distribution. We repeated this analysis for all bootstrapped estimates of OM flux in each stream community. Skewness coefficients exist in the range [-1, 1], where -1 indicates that OM fluxes are skewed perfectly away from a trait and 1 indicates that higher relative fluxes are perfectly associated with higher trait values. To determine if the skewness of fluxes with *M* and *P:B* was related to mean annual stream temperature, we use bootstrapped beta regression with a simple transformation, $(Sk_{flux} + 1)/2$, to meet the assumptions of the model and standardize values between 0 and 1. Model coefficients were back-transformed to estimate effect sizes.

>>>>

---
title: "Environmental warming increases the importance of high-turnover energy channels in stream food webs"
author:
  - James R. Junker:
      email: jrjunker@mtu.edu
      institute: [msu, glrc]
      correspondence: true
  - name: Wyatt F. Cross
    institute: msu
  - name: James M. Hood
    institute: osu
  - name: Jonathan P. Benstead
    institute: bama
  - name: Alexander D. Huryn
    institute: bama
  - name: Daniel Nelson
    institute: usu
  - name: Jón S. Ólafsson
    institute: veidi
  - name: Gísli M. Gíslason
    institute: reyk
institute:
  - msu: Department of Ecology, Montana State University, Bozeman, MT 59717, USA
  - glrc: Current `address:` Great Lakes Research Center, Michigan Technological University, Houghton, MI 49931, USA
  - osu: Department of Evolution, Ecology, and Organismal Biology, The Ohio State University, Translational Data Analytics Institute, The Aquatic Ecology Laboratory, Columbus, OH 43212, USA
  - bama: Department of Biological Sciences, University of Alabama, Tuscaloosa, AL  35487, USA
  - usu: National Aquatic Monitoring Center, Department of Watershed Sciences, Utah State University, Logan, UT 84322, USA
  - veidi: Institute of Marine and Freshwater Fisheries, Hafnarfjördur, Iceland
  - reyk: University of Iceland, Institute of Life and Environmental Sciences, Reykjavík, Iceland
output:
   word_document:
      reference_docx: working_docx_template.docx
      toc: no
      pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
      - '--filter=pandoc-crossref.exe'
   pdf_document:
     keep_tex: true
   html_document: default
   pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
      - '--filter=pandoc-crossref.exe'
header-includes:
- \usepackage{lineno}
- \usepackage{amsmath}
- \numberwithin{equation}
- \usepackage{indentfirst}
- \linenumbers
indent: true
linestretch: 1
bibliography: refs.bib
link-citations: no
link-color: grey
csl: ecology.csl
---

@@

Warming temperatures are altering communities and trophic networks globally. While the overall influence of warming on food webs is often context-dependent, increasing temperatures are predicted to change communities in two fundamental ways: 1) by reducing average body size and 2) by increasing individual metabolic rates. These warming-induced changes have the potential to influence the distribution of food web fluxes, food web stability, and the relative importance of ecological and stochastic processes shaping community assembly. Here, we quantified patterns and the relative distribution of organic matter fluxes through stream food webs spanning a natural ~25$^\circ$C temperature gradient. We then related these patterns to species- and community-level trait distributions of mean body size and population biomass turnover (*P:B*) within and across ecosystems. We predicted that 1) communities in warmer streams would skew towards smaller body size and higher *P:B* and 2) organic matter fluxes within warmer communities would increasingly skew towards smaller, higher *P:B* populations. Across the temperature gradient, communities in warmer streams were on average composed of smaller-bodied and higher *P:B* populations. Additionally, material fluxes *within* warmer streams were increasingly skewed towards high-turnover populations, demonstrating that warming can restructure organic matter fluxes in both an absolute and relative sense. With warming, the relative distribution of organic matter fluxes also appeared to be increasingly 'non-random', suggesting the stronger selection for traits driving high turnover with increasing temperature. Our study suggests that a warming world will be 'smaller and faster', and that warming has the potential to shape pathways of energy and material flows in food webs in ways that may be predictable.

@@

<!--- warming and effects in food webs --->
<!--- Speeding up of food web: Species levels--->
Increasing global temperatures are influencing the provision and maintenance of ecosystem services by modifying the network of species interactions that underpin ecosystem functions [@brose2012; @woodward2010; @deruiter1995; @thompson2012]. The effects of warming permeate across levels of biological organization from its control on individual metabolic rates [@gillooly2001; @brown2004] and biological activity (e.g., attack rate, handling time, growth rates, etc.; @dell2014, to broad-scale shifts in community assembly and structure [@saito2021; @gibert2019a; @nelson2017a]. Across global climate gradients, such temperature-induced changes have the potential to mediate food web stability [@baiser2019] by altering the acquisition and allocation of resources [@zhang2017] and the magnitude and relative distribution of energy fluxes in food webs [@may1972; @mccann1998; @barnes2018].

Understanding how warming alters energy fluxes in food webs requires information about how temperature modifies connections between ecosystem structure and function. Such connections hinge upon how warming influences functional trait distributions and how functional traits translate to energy demands [@loreau2001; @norberg2001; @norberg2004]. Additionally, warming may modify both deterministic [e.g., environmental/niche filtering, @whittaker1962] and stochastic [e.g., neutral theory, @hubbell2001] community assembly processes [@saito2021], leading to many possible relationships between warming and species abundance and trait distributions. For example, although temperature appears to have minimal control on species richness across taxonomic groups [e.g., @bastazini2021], warming may have profound effects on the dominance structure (i.e., evenness) of ecological communities, often by reducing community evenness and favoring a few well-adapted species [@hillebrand2008]. This strong environmental filtering is likely to skew trait distributions in natural communities [@therriault1999]. However, the relative distribution of traits can be modified further by additional processes beyond environmental filtering that alter species relative abundance distributions [e.g., species interactions, @therriault1999;\; demographic stochasticity, @hubbell2001]. These additional processes can exaggerate or counter any skew in trait distributions imparted through environmental filtering and thereby modify the relationship between species' traits and the absolute and relative energy demands in food webs.

<!--- Connect traits to the 'speed' of ecosystem energy fluxes--->
Body size is a fundamental trait that is strongly influenced by temperature and has great potential to influence energy flux [@atkinson1994; @gardner2011; @daufresne2009]. Body size reductions arise from multiple mechanisms acting at different levels of organization, including the increased relative abundance of smaller species within warmer communities [@bergmann1848] or smaller individuals in warmer populations [@james1970], or reduced absolute body size of warmer individuals [@atkinson1994]. These temperature-size relationships are likely to influence the distribution of body sizes both across and within ecosystems, such that warmer communities should contain smaller species and individuals, on average. These relationships can have important implications for ecosystems patterns and processes that control energy and material fluxes in ecosystems because body size is a strong determinant of species life-history patterns [@peters1983; @zeuss2017; @altermatt2010; @nelson2020b], as well as developmental [@angilletta2004] and metabolic rates [@brown2004; @gillooly2001].

<!--- Write the main goals of the paragraph--->
Metabolic rate, in addition to being controlled by body size, is another key trait that is influenced by temperature and tied to variation in energy flux through food webs [@gillooly2001]. Warming can influence metabolic rates indirectly through reductions in body size, as well as directly through its effects on subcellular kinetics [@bideault2019; @osmond2017]. These processes have been shown to modify ecosystem patterns through changes in population abundance [@bernhardt2018], consumer-resource interactions [@bideault2019], and food web structure [@gibert2019a]. Importantly, the kinetic effects of temperature are tied to many biological processes [@dell2014], among them growth rate [@gillooly2001], developmental rate [@zuo2012; @nelson2020b], voltinism [@zeuss2017], and population turnover rate [@huryn2007; @brown2004]. Taken together, these changes suggest that warming should lead to a smaller, faster world in which ecosystem processes 'speed up' through the effects of smaller body size and higher turnover rates, with a potentially strong imprint on food web structure and energy fluxes [@gibert2019a].

We quantified patterns of organic matter flux in stream food webs across a natural temperature gradient (\~5--28$^\circ$C) in southwestern Iceland. Previous research in these streams has shown a strong positive effect of warming on primary production both among streams [@demars2011; @padfield2017] and seasonally within streams [@ogorman2012; @hood2018]. Invertebrates in these streams rely on autotrophs [@ogorman2012; @nelson2020], thus the dynamics of primary production functions as a strong control on energy flow through consumers [@junker2020]. We therefore predicted that annual organic matter fluxes to consumers would increase with stream temperature, mirroring patterns in resource availability and consumer energy demand. We also hypothesized that temperature would act as a principle environmental filter on community assembly and organic matter fluxes by favoring 'fast' life-history traits associated with small-bodied organisms. Specifically, we predicted that warming temperatures among streams would lead to a decrease in average body size and an increase in average biomass turnover (i.e., *P:B* ratio) of populations. We also predicted that *within* communities, organic matter fluxes would be dominated by small-bodied and high *P:B* taxa at higher temperatures and that these patterns would not arise by random change [i.e., a 'selection effect' *sensu* @huston1997]. Our results should help refine general predictions about how ongoing climate warming, and its influence on key traits, is likely to shape energy flux through animal food webs.

@@

We studied six streams within the Hengill geothermal field of southwestern Iceland (64$^\circ$ 03'N 021$^\circ$ 18'W) that varied in mean annual temperature from ~5 to 30$^\circ$C. Hengill is
characterized by indirect geothermal heating of groundwater [@arnason1969], leading to natural variability in stream temperatures (4.5--54.0 $^\circ$C), but similar solute
chemistries [@friberg2009]. These conditions create a "natural laboratory" for isolating the effects of temperature on ecosystem processes [@ogorman2014; @nelson2017a]. We selected streams to maximize the temperature range, while minimizing differences in the structural aspects of the primary producer community [@junker2020a]. In each stream, we measured temperature and water depth every 15 min from July 2010 through August 2012 (U20-001-01 water-level logger, Onset Computer Corp., Pocasset, MA, USA). Light availability in the watershed was measured every 15 min from atmospheric stations (HOBO pendant temperature/light UA-002-64, Onset Computer Corp., Pocasset, MA, USA).

@@

We sampled macroinvertebrate communities approximately monthly in six streams, four from July 2011 to August 2012 and two from October 2010 to October 2011 (*n* = 6 streams). The two streams sampled from 2010--2011 were part of a separate warming manipulation during the reference (i.e., non-manipulated) year [@nelson2017; @nelson2017a]. Inter-annual comparisons of primary and secondary production in previous studies showed minimal differences among years in unmanipulated streams, suggesting that combining data from different years would not significantly bias our results [@nelson2017; @hood2018]. We collected five Surber samples (0.023 m^2^, 250-$\mu$m mesh) from randomly selected locations within each stream. Within the sampler, inorganic substrates were disturbed to \~10 cm depth and invertebrates and
organic matter were removed from stones with a brush. Samples were then preserved with 5% formaldehyde until laboratory analysis. In the laboratory, we split samples into coarse
(\>1 mm) and fine (\>250 $\mu$m--\<1 mm) fractions using nested sieves and then removed invertebrates from each fraction under a dissecting microscope (10--15$\times$ magnification).
For particularly large samples, fine fractions were sub-sampled (1/2--1/16th) using a modified Folsom plankton splitter prior to removal of invertebrates. Macroinvertebrates were identified to the lowest practical taxonomic level (usually genus) with taxonomic keys [@peterson1977; @merritt2008; @andersen2013]. Taxon-specific abundance and biomass were scaled to a per-meter basis by dividing by the Surber sampler area.

@@

Daily secondary production of invertebrate taxa was calculated using the instantaneous growth rate method [IGR, @benke2017]. Growth rates were determined using taxon-appropriate
approaches described in Junker and others [-@junker2020]. Briefly, growth rates of common taxa (e.g., Chironomidae spp., *Radix balthica*, etc.) were determined using *in situ* chambers [@huryn1986].
Multiple individuals (*n* = 5--15) within small size categories (\~1-mm length range) were photographed next to a field micrometer, placed in the stream within pre-conditioned
chambers for 7--15 days, and removed and photographed. Individual lengths were measured from field pictures using image analysis software [@schindelin2012], and body
lengths were converted to mass (mg ash-free dry mass [AFDM]) using published length-mass regressions [@benke1999; @ogorman2012; @hannesdottir2013]. Growth rates (*g*, d^-1^) were
calculated by the changes in mean body size (*M*) over a given time interval (*t*) with the following equation:

@@

Variability in growth rates was estimated by bootstrapping through repeated resampling of individual lengths with replacement (*n* = 1000). For taxa that exhibit synchronous growth and development (e.g., Simuliidae spp., some Chironomidae spp., etc.), we examined temporal changes in length-frequency distributions and calculated growth rates and uncertainty using a bootstrap technique similar to that described in Benke and Huryn [-@benke2017]. Individual lengths were converted to mass (mg AFDM) using published length-mass regression cited above, and size-frequency histograms were visually inspected for directional changes in body size through time. For each date, size-frequency distributions were resampled with replacement and growth rates estimated from equation 1. We prevented the calculation of negative growth rates by requiring $M_{t + \Delta t}$ \> $M_t$. If this condition was not met after 10,000 resamplings, a minimum growth rate of 0.001 was used. To estimate growth rates of taxa for which growth could not be estimated empirically, we developed stream-specific growth rate models by constructing multivariate linear regressions of empirical growth rates against body size and temperature. To estimate uncertainty in production of each taxon, we used a bootstrapping technique that resampled measured growth rates, in addition to abundance and size distributions from individual samples. For each iteration, size-specific growth rates were multiplied by mean
interval biomass for each size class and the number of days between sample dates to estimate size class-specific production. For each interval, size classes were summed for each taxon to calculate total population-level production. Intervals were summed to estimate annual secondary production (g AFDM m^-2^ y^-1^).

@@

Organic matter fluxes through the community were calculated using the trophic basis of production method [TBP\; @benke1980]. Taxon-specific secondary production estimates were combined with diet proportions (see Supplemental Materials), diet-specific assimilation efficiencies, *AE~i~*, and net production efficiencies, *NPE* [e.g., @mccullough1975], to estimate consumption of organic matter. For each food
category, *i*, diet proportions were multiplied by the gross growth efficiency ($GGE_{i} = AE_{i} * NPE$) to calculate the relative production attributable to each food
category. The relative production from each food type was then multiplied by the interval-level production and finally divided by $GGE_{i}$ to estimate consumption of organic matter from each food category by each taxon [@benke1980]. Consumption was calculated for each taxon across sampling intervals (typically \~1 month). Total interval consumption was calculated by summing across all taxa, while annual consumption was calculated by summing across all taxa and intervals. Variability in consumption estimates was estimated through a Monte Carlo approach, wherein bootstrapped vectors of secondary production for each taxon (see *Secondary production* methods above) were resampled and consumption estimated with the TBP method using modeled diet proportions (see *Diet analysis* in Supplemental Materials), diet-specific assimilation efficiencies, and net production efficiency. Variability in *AE~i~* was incorporated by resampling values from beta distributions fit to median and 2.5% and 97.5% percentiles for each diet item: diatoms = 0.30 (95% percentile interval (PI): 0.24-0.36), filamentous and green algae = 0.30 (95% PI: 0.24-0.36), cyanobacteria = 0.10 (95% PI: 0.08-0.12), amorphous detritus = 0.10 (95% PI: 0.08-0.12), vascular and non-vascular plants (bryophytes) = 0.1 (95% PI: 0.08-0.12), and animal material = 0.7 (95% PI: 0.56-0.84)[@welch1968; @cross2007; @cross2011; @benke1980; @benke1997]. Variability in *NPE* was incorporated by resampling values from an assumed beta distribution with median *NPE* = 0.45 (95% PI = 0.4-0.5). Beta distributions were fit in R [@rcoreteam2022] using the 'get.beta.par()' function within the *rriskDistributions* package [@belgorodski2017].

@@

### Evenness among taxa

To visualize and quantify the evenness of OM fluxes among taxa within each stream, we constructed Lorenz curves [@lorenz1905] on rank-ordered OM fluxes, such that in a community with $S$ species and the relative OM flux of species *i*, $p_i$, is ordered $p_1 \leq p_2 \leq ... p_S$. The Lorenz curve shows how a value, in this case OM flux, accumulates with increasing cumulative proportion of taxa. In a community with an equal distribution of OM flux among taxa, the Lorenz curve is simply a straight diagonal line. Deviation from equality was calculated as the Gini coefficient [@gini1921], normalized for differences in $S$ among streams, $G^*$ [@solomon1975; @chao2019]:

@@

where $G^*$ represents an index of relative evenness of OM fluxes bounded between zero and one; a value of one represents a community with equal proportion of total community OM flux for all species ($1/S$), and a value of zero represents a community in which the total OM flux is attributed to a single taxon.

@@

We predicted that warming would favor taxa with smaller body size and higher population turnover and therefore OM fluxes would be skewed towards small body size (*M*) and higher *P:B* across and within communities. To assess the potential effects of environmental filtering across communities, we used bootstrapped linear regressions to quantify the relationship between mean annual temperature ($^\circ$C) and 1) mean population body size ($\overline{M}$) and 2) biomass turnover ($\overline{P:B}$) of each stream community. Here, `r format(temperature_stats[['n_boot']], big.mark = ",", scientific = FALSE)` values of population *M* or *P:B* were resampled with replacement from population secondary production vectors (see *Secondary Production* above) for each taxon within each stream. The mean of all populations within a stream was calculated and a linear model was fit between *log~e~*-transformed $\overline{M}$ or $\overline{P:B}$ and mean annual temperature. Response variables were transformed to meet the assumption of normally distributed residual variation.

To quantify how the distribution of organic matter fluxes were modified *within* a community and whether this modification was related to temperature, we assessed the extent to which the relative OM fluxes among taxa were skewed towards populations with lower or higher relative *M* or *P:B*. To do this, we ordered taxa based on within-stream rankings of annual population traits (i.e., *M*, *P:B*) and then calculated a measure of skewness, $Sk_{flux}$, based on quartiles of the distribution of OM fluxes in relation to taxon traits as:

@@

where, $f_{Qx}$, is the cumulative flux at some quantile, $Qx$, of the community trait distribution. We repeated this analysis for all estimates of OM flux used to calculate *M* and *P:B* in each stream community. Skewness coefficients exist in the range [-1, 1], where -1 indicates that OM fluxes are skewed perfectly away from a trait and 1 indicates that higher relative fluxes are perfectly associated with higher trait values. To determine if the skewness of fluxes with *M* and *P:B* was related to mean annual stream temperature, we use bootstrapped beta regression with a simple transformation, $(Sk_{flux} + 1)/2$, to meet the assumptions of the model and standardize values between 0 and 1. Model coefficients were back-transformed to estimate effect sizes.
